Does at-issueness predict projection? Further investigations of the Gradient Projection Principle

### 2023-03-14 writing/SuB-abstract/SuB-abstract.tex
section 2 updates

\vspace{-.7\baselineskip}
\noindent 
	We present experimental data that the projection of the content of attitude complements \textbf{(i)} varies between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
	%
	The observed variability is not captured by existing theoretical accounts of projection
	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), but suggests that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes. Instead, our results reveal a more nuanced picture of lexical semantic and pragmatic properties, which raises important questions for future research on projection.
	%

\noindent {\bf Projection across entailment-cancelling operators.}
	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.

@@

	\vspace{-.5\baselineskip}
	\noindent  Current research only rarely examines differences in projectivity across these four operators, with conflicting findings.
	%
	\citet{smith_relationship_2014} found that that the content of non-restrictive relative clauses and the preparatory content of \emph{win} was more projective under conditionals than negation, whereas the projective content of epithets and the clausal complement of \emph{know} showed the opposite pattern.
	%
	\citet{karttunen_observations_1971} proposed that the CC of English factive predicates (e.g. \emph{be annoyed, regret}) projects across all four operators, whereas that of semi-factives (e.g. \emph{discover, realize, see, notice}) only projects across negation.
	%
	In contrast, \posscite{sieker_projective_2022} investigation of the projection of the CC of German attitude predicates found higher projectivity under negation than other operators (e.g. modals), but no interaction with predicate type. In particular, they did not find a difference between factive and semi-factive verbs, nor did they replicate \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}).
	%
	The divergent results raise the question of whether they are due to variation between English and German or methodological differences.

\noindent
{\bfseries Experiment.}
	%
	We conducted a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}, using the `certain that' diagnostic (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020,sieker_projective_2022}). We applied this diagnostic to the contents of the complements of 20 English clause-embedding predicates, including purported factive predicates (e.g., \emph{be annoyed, know, reveal}) and purported semi-factive predicates (e.g., \emph{discover, see}),
	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).

\noindent 
{\bf Methods and Expectations.}
	Projection of the CC of the 20 attitude predicates was measured in four sets of experiments: The predicates were embedded in polar questions in Exps.~1, under negation (Exps.~2), under {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4). (Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block. We focus on the projection ratings here.) In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?). Participants gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other. Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator. We analyze data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform. Based on Karttunen's generalization we would expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and lower ratings under the other operators.

	% the projection of the complement clause (CC) of 20 attitude predicates was measured in four sets of experiments: polar questions (Exps. 1), negation (Exps. 2), perhaps (Exps. 3), and conditional antecedents (Exps. 4). Each set had three experiments, using different measures. Participants judged whether the speaker was certain of the CC and responded on a slider from 0 to 1. Each participant saw all 20 attitude predicates paired with a unique content under one operator. Data was analyzed from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon’s MT platform. Factive predicates were expected to receive high projection ratings under all four operators, while semi-factive predicates were expected to have high ratings under negation and lower ratings under the other operators.

\noindent
{\bf Results and Analysis.} 
	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
	%
	We find \textbf{(i)} differences in projection between operators, even if we aggregate across predicates, and \textbf{(ii)} different differences (interactions) between operators and predicates. \textbf{(i)} the differences between operators, when pooling predicates are small, but signifgicant. This is supported by a linear mixed effects model, with fixed effect: \texttt{operator}, reference level: \texttt{question}; random effect: \texttt{participant} intercepts; Model output: \texttt{question} (baseline): $0.515$ (Maximum Likelihood Estimate), $0.006$ Standard Error; lower projectivity for \texttt{modal}: $-0.04$ MLE, $0.009$ SE, $|t| = 4.55$, $p < 0.001$; \texttt{negation}: $-0.027$ MLE, $0.006$ SE, $|t| = 4.3$, $p < 0.001$; higher projectivity for \texttt{conditional}: $0.044$ MLE, $0.009$ SE, $|t| = 5.07$, $p < 0.001$.\\

	\noindent
	{\bfseries LH: } I still wonder if it might be enough to appeal to a main effect of operator in a full model (with both predictor variables) to support point (i). I know it's not as thourough, but shows that this effect is present at least in the baseline, and saves space.\\

	\noindent \textbf{(ii)} differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals. These results (supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}).\\

	\noindent
	{\bfseries LH: } Here it may be enough to report interaction terms in a full model? Unless we explicitly want to make the point against the factive / semi-factive distinction.\\
	% and significant interaction effects in these models (e.g. the first model has $n$?? significant interaction terms with $|t| > x$, and $p < y$...)

	\noindent Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.\\

@@

\noindent {\bf Discussion:  Empirical implications and lexical properties.} \\
	%
	% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates. However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions. \citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive). Here, it is also not obvious how exactly this task relates to projection. \citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC. The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
	%
	\noindent
	{\bfseries LH: } Here, JT said before that we should discuss cross-linguistic implications re Sieker \& Solstad. What would our conlusion be here?.\\

	The predicate / operator interactions we find for the selected English attitude predicates are not in line with \cite{sieker_projective_2022}, who find no such effects for the selected German attitudes, but used the same task as the study presented here. This difference in findings indicates potential cross-linguistic differences in projection variability, which provides an avenue for future research.\\

@@

\begin{table}[h]
	\vspace{-1\baselineskip}\phantom{.}
	\centering
	% \hspace{-1.2em}
	\begin{tabular}{p{.65\linewidth} p{.3\linewidth}}
		\scalebox{.72}{
			\begin{tabular}{llrrrr}
				Model & & Estimate & Std. Error & t-value\\
				\midrule
				\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
				& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
				& operator: modal & -0.16 & 0.02  & -10.04 & *** \\
				& operator: question & 0.02 & 0.01  & 1.74 & n.s. \\
				\midrule
				\#2 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 &  69.24 & *** \\
				&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
				&		operator: modal  & -0.14 & 0.02  & -9.2 & *** \\
				&		operator: question & 0.08 & 0.01  & 5.72 & *** \\
				\midrule
				\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
				& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
				&		operator: modal & -0.06 & 0.02  & -3.6 & *** \\
				&		operator: question & 0.1 & 0.01  & 7.07 & *** \\

				\bottomrule
			\end{tabular}
		}
		&
		\parbox{\linewidth}{\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC. Models were fit with \texttt{lme4, lmertest} in \texttt{R}.}\label{t:models}}

@@

\vspace{-.4\baselineskip}
\noindent \normalsize Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).

@@

% \begin{table}[ht]
% 	\caption{Model output w \texttt{be\_annoyed/n} as baseline}
% 	\label{tab:annoyed-n}
% 	\centering

% 	\begin{tabular}{lrrrrr}
% 		\toprule
% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
% 		\midrule
% 		(Intercept) & 0.8669958 & 0.0114402 & 1479.204 & 75.7850523 & 0.0000000\\
% 		opc & -0.1155850 & 0.0156641 & 26227.368 & -7.3789836 & 0.0000000\\
% 		opm & -0.1564480 & 0.0155858 & 26288.139 & -10.0378702 & 0.0000000\\
% 		opq & 0.0249138 & 0.0143090 & 56826.915 & 1.7411232 & 0.0816674\\
% 		verbsuggest & -0.6162683 & 0.0134941 & 54372.895 & -45.6694112 & 0.0000000\\
% 		\addlinespace
% 		verbacknowledge & -0.2204417 & 0.0134948 & 54373.777 & -16.3352724 & 0.0000000\\
% 		verbadmit & -0.3402430 & 0.0134933 & 54371.874 & -25.2156750 & 0.0000000\\
% 		verbannounce & -0.3607504 & 0.0134931 & 54371.547 & -26.7359958 & 0.0000000\\
% 		verbbe\_right & -0.6638924 & 0.0134936 & 54372.293 & -49.2003974 & 0.0000000\\
% 		verbconfess & -0.3681695 & 0.0134949 & 54373.904 & -27.2820517 & 0.0000000\\
% 		\addlinespace
% 		verbconfirm & -0.6338363 & 0.0134931 & 54371.657 & -46.9746939 & 0.0000000\\
% 		verbdemonstrate & -0.5307026 & 0.0134935 & 54372.157 & -39.3301466 & 0.0000000\\
% 		verbdiscover & -0.1864716 & 0.0134942 & 54372.992 & -13.8186520 & 0.0000000\\
% 		verbestablish & -0.5542561 & 0.0134933 & 54371.853 & -41.0764079 & 0.0000000\\
% 		verbhear & -0.1956794 & 0.0134938 & 54372.473 & -14.5014554 & 0.0000000\\
% 		\addlinespace
% 		verbinform & -0.1374428 & 0.0134945 & 54373.345 & -10.1851106 & 0.0000000\\
% 		verbknow & -0.0748953 & 0.0134936 & 54372.288 & -5.5504197 & 0.0000000\\
% 		verbpretend & -0.5049533 & 0.0134944 & 54373.218 & -37.4195409 & 0.0000000\\
% 		verbprove & -0.6364950 & 0.0134935 & 54372.138 & -47.1704273 & 0.0000000\\
% 		verbreveal & -0.3057814 & 0.0134937 & 54372.418 & -22.6609884 & 0.0000000\\
% 		\addlinespace
% 		verbsay & -0.5419725 & 0.0134925 & 54370.770 & -40.1685159 & 0.0000000\\
% 		verbsee & -0.1827512 & 0.0134932 & 54371.791 & -13.5438968 & 0.0000000\\
% 		verbthink & -0.5101198 & 0.0134930 & 54371.491 & -37.8062103 & 0.0000000\\
% 		opc:verbsuggest & 0.1594325 & 0.0190634 & 54372.626 & 8.3632939 & 0.0000000\\
% 		opm:verbsuggest & 0.1766113 & 0.0189718 & 54372.046 & 9.3091498 & 0.0000000\\
% 		\addlinespace
% 		opq:verbsuggest & -0.0737648 & 0.0192906 & 54372.717 & -3.8238781 & 0.0001315\\
% 		opc:verbacknowledge & 0.1945569 & 0.0190630 & 54372.333 & 10.2059764 & 0.0000000\\
% 		opm:verbacknowledge & 0.1100553 & 0.0189729 & 54373.070 & 5.8006461 & 0.0000000\\
% 		opq:verbacknowledge & 0.0672935 & 0.0192909 & 54372.991 & 3.4883599 & 0.0004864\\
% 		opc:verbadmit & 0.2803349 & 0.0190626 & 54371.933 & 14.7060204 & 0.0000000\\
% 		\addlinespace
% 		opm:verbadmit & 0.1810708 & 0.0189717 & 54371.994 & 9.5442376 & 0.0000000\\
% 		opq:verbadmit & 0.0440050 & 0.0192896 & 54371.832 & 2.2812856 & 0.0225354\\
% 		opc:verbannounce & 0.2503929 & 0.0190639 & 54373.132 & 13.1343759 & 0.0000000\\
% 		opm:verbannounce & 0.1633026 & 0.0189721 & 54372.338 & 8.6075049 & 0.0000000\\
% 		opq:verbannounce & 0.0017532 & 0.0192899 & 54372.121 & 0.0908892 & 0.9275810\\
% 		\addlinespace
% 		opc:verbbe\_right & 0.1424317 & 0.0190623 & 54371.698 & 7.4718895 & 0.0000000\\
% 		opm:verbbe\_right & 0.2582399 & 0.0189717 & 54371.967 & 13.6118412 & 0.0000000\\
% 		opq:verbbe\_right & -0.0776846 & 0.0192897 & 54371.974 & -4.0272512 & 0.0000565\\
% 		opc:verbconfess & 0.3111230 & 0.0190643 & 54373.455 & 16.3196605 & 0.0000000\\
% 		opm:verbconfess & 0.1837114 & 0.0189741 & 54374.125 & 9.6821973 & 0.0000000\\
% 		\addlinespace
% 		opq:verbconfess & 0.0393640 & 0.0192890 & 54371.320 & 2.0407459 & 0.0412809\\
% 		opc:verbconfirm & 0.2963833 & 0.0190629 & 54372.237 & 15.5476205 & 0.0000000\\
% 		opm:verbconfirm & 0.3663333 & 0.0189721 & 54372.324 & 19.3090505 & 0.0000000\\
% 		opq:verbconfirm & 0.0312870 & 0.0192902 & 54372.377 & 1.6219108 & 0.1048282\\
% 		opc:verbdemonstrate & 0.2854366 & 0.0190647 & 54373.829 & 14.9719656 & 0.0000000\\
% 		\addlinespace
% 		opm:verbdemonstrate & 0.2783816 & 0.0189716 & 54371.830 & 14.6736292 & 0.0000000\\
% 		opq:verbdemonstrate & 0.0753554 & 0.0192892 & 54371.520 & 3.9066023 & 0.0000937\\
% 		opc:verbdiscover & 0.2270099 & 0.0190632 & 54372.450 & 11.9082962 & 0.0000000\\
% 		opm:verbdiscover & 0.1002424 & 0.0189726 & 54372.795 & 5.2835252 & 0.0000001\\
% 		opq:verbdiscover & 0.0762282 & 0.0192898 & 54372.014 & 3.9517423 & 0.0000777\\
% 		\addlinespace
% 		opc:verbestablish & 0.2570817 & 0.0190633 & 54372.539 & 13.4857080 & 0.0000000\\
% 		opm:verbestablish & 0.2879581 & 0.0189714 & 54371.677 & 15.1785426 & 0.0000000\\
% 		opq:verbestablish & 0.0061604 & 0.0192893 & 54371.583 & 0.3193687 & 0.7494481\\
% 		opc:verbhear & 0.2016309 & 0.0190630 & 54372.273 & 10.5770983 & 0.0000000\\
% 		opm:verbhear & -0.0046313 & 0.0189714 & 54371.681 & -0.2441216 & 0.8071376\\
% 		\addlinespace
% 		opq:verbhear & 0.0756021 & 0.0192902 & 54372.373 & 3.9192008 & 0.0000890\\
% 		opc:verbinform & 0.1278740 & 0.0190645 & 54373.631 & 6.7074362 & 0.0000000\\
% 		opm:verbinform & 0.0141486 & 0.0189741 & 54374.122 & 0.7456761 & 0.4558664\\
% 		opq:verbinform & 0.0587564 & 0.0192913 & 54373.322 & 3.0457537 & 0.0023221\\
% 		opc:verbknow & 0.1142699 & 0.0190629 & 54372.206 & 5.9943629 & 0.0000000\\
% 		\addlinespace
% 		opm:verbknow & 0.0130032 & 0.0189709 & 54371.181 & 0.6854301 & 0.4930755\\
% 		opq:verbknow & 0.0569598 & 0.0192897 & 54371.975 & 2.9528579 & 0.0031498\\
% 		opc:verbpretend & -0.0902047 & 0.0190640 & 54373.155 & -4.7316842 & 0.0000022\\
% 		opm:verbpretend & -0.0194245 & 0.0189738 & 54373.822 & -1.0237558 & 0.3059552\\
% 		opq:verbpretend & -0.2196823 & 0.0192900 & 54372.250 & -11.3883795 & 0.0000000\\
% 		\addlinespace
% 		opc:verbprove & 0.2579585 & 0.0190627 & 54372.031 & 13.5320999 & 0.0000000\\
% 		opm:verbprove & 0.3580525 & 0.0189720 & 54372.223 & 18.8726898 & 0.0000000\\
% 		opq:verbprove & 0.0230805 & 0.0192908 & 54372.899 & 1.1964509 & 0.2315259\\
% 		opc:verbreveal & 0.2902650 & 0.0190644 & 54373.511 & 15.2255216 & 0.0000000\\
% 		opm:verbreveal & 0.1767476 & 0.0189723 & 54372.513 & 9.3160804 & 0.0000000\\
% 		\addlinespace
% 		opq:verbreveal & 0.0665887 & 0.0192906 & 54372.717 & 3.4518763 & 0.0005571\\
% 		opc:verbsay & 0.1985351 & 0.0190618 & 54371.193 & 10.4153412 & 0.0000000\\
% 		opm:verbsay & 0.1451974 & 0.0189723 & 54372.495 & 7.6531273 & 0.0000000\\
% 		opq:verbsay & -0.1815738 & 0.0192889 & 54371.195 & -9.4133926 & 0.0000000\\
% 		opc:verbsee & 0.2175191 & 0.0190629 & 54372.234 & 11.4105822 & 0.0000000\\
% 		\addlinespace
% 		opm:verbsee & 0.0594092 & 0.0189723 & 54372.499 & 3.1313655 & 0.0017409\\
% 		opq:verbsee & 0.1108274 & 0.0192896 & 54371.840 & 5.7454526 & 0.0000000\\
% 		opc:verbthink & 0.0023197 & 0.0190619 & 54371.272 & 0.1216913 & 0.9031440\\
% 		opm:verbthink & 0.0145577 & 0.0189717 & 54371.965 & 0.7673367 & 0.4428847\\
% 		opq:verbthink & -0.1738342 & 0.0192903 & 54372.438 & -9.0115053 & 0.0000000\\
% 		\bottomrule
% 	\end{tabular}
% \end{table}

% \newpage

% \begin{table}[ht]
% 	\caption{Model output w \texttt{discover/n} as baseline}
% 	\label{tab:discover-n}
% 	\centering

% 	\begin{tabular}{lrrrrr}
% 		\toprule
% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
% 		\midrule
% 		(Intercept) & 0.6805242 & 0.0114403 & 1479.217 & 59.4849565 & 0.0000000\\
% 		opc & 0.1114249 & 0.0156639 & 26226.686 & 7.1134644 & 0.0000000\\
% 		opm & -0.0562056 & 0.0155855 & 26286.826 & -3.6062779 & 0.0003112\\
% 		opq & 0.1011420 & 0.0143103 & 56828.295 & 7.0677830 & 0.0000000\\
% 		verbbe\_annoyed & 0.1864716 & 0.0134942 & 54372.998 & 13.8186521 & 0.0000000\\
% 		\addlinespace
% 		verbsuggest & -0.4297967 & 0.0134937 & 54372.361 & -31.8516905 & 0.0000000\\
% 		verbacknowledge & -0.0339702 & 0.0134936 & 54372.308 & -2.5174950 & 0.0118221\\
% 		verbadmit & -0.1537714 & 0.0134942 & 54372.985 & -11.3953857 & 0.0000000\\
% 		verbannounce & -0.1742788 & 0.0134947 & 54373.649 & -12.9145926 & 0.0000000\\
% 		verbbe\_right & -0.4774208 & 0.0134935 & 54372.079 & -35.3816304 & 0.0000000\\
% 		\addlinespace
% 		verbconfess & -0.1816980 & 0.0134936 & 54372.192 & -13.4655376 & 0.0000000\\
% 		verbconfirm & -0.4473648 & 0.0134944 & 54373.311 & -33.1517710 & 0.0000000\\
% 		verbdemonstrate & -0.3442310 & 0.0134928 & 54371.247 & -25.5121517 & 0.0000000\\
% 		verbestablish & -0.3677846 & 0.0134928 & 54371.167 & -27.2579106 & 0.0000000\\
% 		verbhear & -0.0092079 & 0.0134940 & 54372.798 & -0.6823676 & 0.4950095\\
% 		\addlinespace
% 		verbinform & 0.0490288 & 0.0134942 & 54373.066 & 3.6333114 & 0.0002801\\
% 		verbknow & 0.1115762 & 0.0134931 & 54371.668 & 8.2691029 & 0.0000000\\
% 		verbpretend & -0.3184818 & 0.0134943 & 54373.176 & -23.6011440 & 0.0000000\\
% 		verbprove & -0.4500234 & 0.0134934 & 54372.042 & -33.3512833 & 0.0000000\\
% 		verbreveal & -0.1193099 & 0.0134941 & 54372.880 & -8.8416323 & 0.0000000\\
% 		\addlinespace
% 		verbsay & -0.3555009 & 0.0134943 & 54373.168 & -26.3444678 & 0.0000000\\
% 		verbsee & 0.0037204 & 0.0134930 & 54371.450 & 0.2757288 & 0.7827574\\
% 		verbthink & -0.3236483 & 0.0134936 & 54372.197 & -23.9853913 & 0.0000000\\
% 		opc:verbbe\_annoyed & -0.2270099 & 0.0190632 & 54372.453 & -11.9082963 & 0.0000000\\
% 		opm:verbbe\_annoyed & -0.1002424 & 0.0189726 & 54372.798 & -5.2835253 & 0.0000001\\
% 		\addlinespace
% 		opq:verbbe\_annoyed & -0.0762282 & 0.0192898 & 54372.017 & -3.9517423 & 0.0000777\\
% 		opc:verbsuggest & -0.0675773 & 0.0190625 & 54371.851 & -3.5450396 & 0.0003929\\
% 		opm:verbsuggest & 0.0763690 & 0.0189709 & 54371.211 & 4.0255864 & 0.0000569\\
% 		opq:verbsuggest & -0.1499930 & 0.0192910 & 54373.069 & -7.7752981 & 0.0000000\\
% 		opc:verbacknowledge & -0.0324529 & 0.0190617 & 54371.116 & -1.7025196 & 0.0886637\\
% 		\addlinespace
% 		opm:verbacknowledge & 0.0098129 & 0.0189711 & 54371.417 & 0.5172570 & 0.6049789\\
% 		opq:verbacknowledge & -0.0089347 & 0.0192902 & 54372.396 & -0.4631724 & 0.6432426\\
% 		opc:verbadmit & 0.0533251 & 0.0190632 & 54372.508 & 2.7972739 & 0.0051554\\
% 		opm:verbadmit & 0.0808284 & 0.0189728 & 54372.916 & 4.2602367 & 0.0000205\\
% 		opq:verbadmit & -0.0322232 & 0.0192908 & 54372.883 & -1.6703960 & 0.0948468\\
% 		\addlinespace
% 		opc:verbannounce & 0.0233831 & 0.0190642 & 54373.367 & 1.2265427 & 0.2199998\\
% 		opm:verbannounce & 0.0630602 & 0.0189726 & 54372.754 & 3.3237576 & 0.0008887\\
% 		opq:verbannounce & -0.0744750 & 0.0192926 & 54374.465 & -3.8602856 & 0.0001134\\
% 		opc:verbbe\_right & -0.0845782 & 0.0190620 & 54371.360 & -4.4370099 & 0.0000091\\
% 		opm:verbbe\_right & 0.1579976 & 0.0189719 & 54372.163 & 8.3279685 & 0.0000000\\
% 		\addlinespace
% 		opq:verbbe\_right & -0.1539128 & 0.0192901 & 54372.283 & -7.9788617 & 0.0000000\\
% 		opc:verbconfess & 0.0841131 & 0.0190632 & 54372.437 & 4.4123417 & 0.0000102\\
% 		opm:verbconfess & 0.0834691 & 0.0189719 & 54372.142 & 4.3996174 & 0.0000109\\

>>>>

% 1
\vspace{-.7\baselineskip}
\noindent 
	We present experimental data that the projection of the content of attitude complements \textbf{(i)} varies between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
	%
	The observed variability is not captured by existing theoretical accounts of projection
	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), but suggests that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes. Instead, our results reveal a more nuanced picture of lexical semantic and pragmatic properties, which raises important questions for future research on projection.
	%

% 2
\noindent {\bf Projection across entailment-cancelling operators.}
	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.

@@

	\vspace{-.5\baselineskip}
	\noindent  Current research only rarely examines projectivity differences across these operators, with conflicting findings.
	%
	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it only projects across negation.
	%
	Comparing different types of contents, \citet{smith_relationship_2014} found that that the projective content of epithets and the CC of \emph{know} was more projective under negation than conditionals,
	whereas the content of non-restrictive relative clauses and the preparatory content of \emph{win} showed the opposite pattern.
	% whereas the projective contents of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
	%
	Further, \posscite{sieker_projective_2022} investigation of the projection of the CC of German purported factive and semi-factive predicates replicated \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}), but as part of a pattern of overall higher projectivity under negation than other operators. They found no interaction with predicate type, or evidence for the (semi-)factive distinction.

	All of these studies suggest a comparatively high projectivity across negation, but the results diverge about whether they found interactions between operators and different types of contents. To investigate if the different results are due to variation between English and German, differences in contents being tested, or methodological differences, we conducted a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}, using the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive predicates (\emph{be annoyed, know, reveal}) and semi-factives (\emph{discover, see}),
	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}). Our study aimed to replicate the projection variability across operators of various English projective contents reported in Smith and Hall (2014) for English attitude complements.

\noindent 
{\bf Method.}
	Projection of the CC of the 20 attitude predicates was measured in four sets of experiments: The predicates were embedded in polar questions in Exps.\ 1, under negation (Exps.\ 2), under {\em perhaps} (Exps.\ 3), and in conditional antecedents (Exps.\ 4). (Each set contained three experiments using different at-issueness measure in a separate block. Here, we focus on the projection ratings.)
	%
	In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?). Participants gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other. Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator. We analyze data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
	% Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.

\noindent
{\bf Results and Analysis.} 
	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
	%
	We find \textbf{(i)} by-operator differences in projectivity, and \textbf{(ii)} differences in by-operator variability between predicates.
	%
	The generalization \textbf{(i)} is supported by significant main effects of operator in the linear mixed effects models shown in \textbf{Table\ \ref{t:models}}. \textbf{(ii)} is supported by the fact that there are at leats $34$ significant interaction terms in these models (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}). Specifically, the models in \textbf{Table\ \ref{t:models}} illustrate specific differences across the predicates in by-operator variation with different reference levels for \texttt{predicate}: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals, and the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.
	%
	Therefore, we did not replicate the result from \citet{smith_relationship_2014}, that the CC of \emph{know} projects more from negation than conditionals. Given that our findings indicate many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from \posscite{sieker_projective_2022} findings for German, where no such interactions were found.
	%
	Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.

@@

\noindent {\bf Discussion:  Empirical implications and lexical properties.}
	%
	% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates. However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions. \citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive). Here, it is also not obvious how exactly this task relates to projection. \citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC. The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
	%
	% \noindent
	% {\bfseries LH: } Here, JT said before that we should discuss cross-linguistic implications re Sieker \& Solstad. What would our conlusion be here?.\\
	%
	The predicate / operator interactions we find for the selected English attitude predicates are not in line with \cite{sieker_projective_2022}, who find no such effects for the selected German attitudes, but used the same task as the study presented here. This difference in findings indicates potential cross-linguistic differences in projection variability, which provides an avenue for future research.

@@

\begin{table}[h]
	\vspace{-1\baselineskip}\phantom{.}
	\centering
	\hspace{-1.2em}
	\begin{tabular}{p{.663\linewidth} p{.3\linewidth}}
		\scalebox{.77}{
			\begin{tabular}{llrrrr}
				Model & & Estimate & Std. Error & t-value\\
				\midrule
				\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01 & 79.86 & ***\\
				& operator: conditional & -0.12 & 0.02  & -7.36 & ***\\
				& operator: modal & -0.16 & 0.02  & -10.01 & ***\\
				& operator: question & 0.02 & 0.01 & 1.72 & n.s.\\
				\midrule
				\#2 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01 & 62.70 & ***\\
				& operator: conditional & 0.11 & 0.02 & 7.11 & ***\\
				& operator: modal & -0.06 & 0.02 & -3.63 & ***\\
				& operator: question & 0.10 & 0.01 & 7.08 & ***\\
				\midrule
				\#3 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 & 72.97 & ***\\
				& operator: conditional & 0.00 & 0.02 & -0.06 & n.s.\\
				& operator: modal & -0.14 & 0.02 & -9.18 & ***\\
				& operator: question & 0.08 & 0.01 & 5.67 & ***\\
				\bottomrule
			\end{tabular}
		}
		&
		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; with fixed effects of operator, predicate, and their interaction; random effect: participant intercepts. Models were fit with \texttt{lme4, lmertest} in \texttt{R}. All three models also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (Not shown here).}\label{t:models}}

@@

\vspace{-.4\baselineskip}
\noindent \small Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).

@@
